### 2015-09-13 paper-1/main.tex
Most of Michelle's comments are addressed.

%                 Doubly-Stopped Binomial Distribution
%                        Kane and Zelterman
\typeout{}\typeout{}\typeout{}\typeout{}
\typeout{ Two-Stage, Phase II Designs }
\typeout{}\typeout{}
\typeout{ Michael Kane \& Dan Zelterman }
\typeout{}\typeout{}\typeout{}\typeout{}

\documentclass[12pt]{article}         %LaTeX preamble
\parskip   .4em
\jot  2ex
\parindent .3in
\textwidth      6.5in
\oddsidemargin    0in
\textheight 8.75in
\topmargin      -.5in
\baselineskip    26pt
\renewcommand{\baselinestretch}{1.5}
\renewcommand{\arraystretch}{.75}     % vertical spacing in tables
\usepackage{graphicx}                 % Needed to import pdf files
\usepackage{adjustbox}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{url}
\begin{document}

@@

\vspace*{1.25in}
\begin{tabular}{l} 
 {\bf Michael Kane${}^*$} \\
 {\bf Daniel Zelterman}  \\[.1in]

@@

In the Simon 2-stage Phase II clinical trial [\ref{Simon 1989}], a small 
number of patients are enrolled onto a therapeutic 
protocol for an initial stage. Fore each patient, we assess either success 
or failure according to a 
clinically useful criteria. Success might include tumor response to treatment
or perhaps that the response is maintained for a minimum specified duration.
In addition to failing to achieve the primary endpoint, treatment failure 
might include adverse reaction to the drug or withdrawal due to toxicity. We 
use a large number of these early successes as evidence for whether to treat
additional patients in a second stage. A statistically small number of 
successes in Stage I would lead us to terminate the trial without entering 
Stage II.

@@

% ---------------------------  2. Notation ----------------------
\section   {Notation and summary of distributional results}
\label{notation.section}
% --------------------------------------------------------------

Let $\,B_1, B_2, \ldots \,$ denote a sequence of independent, identically 
distributed, Bernoulli random variables with $\mathbb{P}\{B_i=1\}=p$, for 
probability parameter $0\leq p \leq 1$. In the clinical trial setting 
$\,B_i = 0$ corresponds to a successful enrollment. 
Let $s$ and $t$ be positive integers. 
Define the stopped negative binomial (SNB) random variable $Y$ as the smallest 
integer value such that $\,\{B_1, \ldots , B_Y\}\,$ contains {\em either} 
$\,s\,$ successes or else $\,t\,$ failures. The distribution of $\,Y\,$ has 
support on integer values in the range 
\begin{equation*}                                     %   (3)
     \min(s,t) \leq \; Y \;\leq s+t-1  \label{range.y.eq}
\end{equation*}
and it is distributed as
\begin{equation} \label{eqn:pmf}
\mathbb{P}\{Y=k\} = S(k, p, s)  \{s \leq k\} + T(k, p, t) \{t \leq k \} 
\end{equation}
where
\begin{equation} \label{eqn:N}
S(k, p, s) = {k-1 \choose s-1} p^s (1-p)^{k-s} 
\end{equation}
is the negative binomial probability of $s$ successes and
\begin{equation} \label{eqn:R}
T(k, p, t) = {k-1 \choose k-t} p^{k-t} (1-p)^t
\end{equation}
is the negative binomial probability of $t$ failures.

@@

To illustrate that (\ref{eqn:pmf}) is the distribution of $Y$, consider the 
process $\mathbf{X} = \{X(k) : k \in \mathbb{Z}_{\ge 0} \}$ with $X(0)=0$ and
\begin{equation*} \label{eqn:proc}
X_{k+1} = X_k + B_{k+1} \{ k-t < X_k < s \}
\end{equation*}
where $B_{k+1}$ is once again distributed as Bernoulli($p$) for all $k$ in the domain of $\mathbf{X}$. The process can be conceptualized as a series of coin flips that stops when either $s$ heads or $t$ tails are reached. At each step a coin is flipped. If it is heads, the process advances one diagonally in the positive horizontal and vertical direction. Otherwise, it advances in the positive horizontal direction only. The process stops, $X_k$'s value does not change, after either $X_k = s$ or $X_k = k-t$. One example is given in Fig. \ref{fig:kane_viz}.

\begin{prop}
The distribution of the stopping time $\{k: s \leq X_k \leq k-t \}$ is equal to the SNB and is given in (\ref{eqn:pmf}). 
\end{prop}
\begin{proof}
The proof will proceed in two parts. First, a combinatorial justification will be given for the probability mass value on each element of the support. Second, it will be shown that the sum of the masses over the support sums to one.

@@

Let $I_p(s, t)$ be the {\em regularized incomplete beta function}, which is 
also one minus the c.d.f. of the negative binomial distribution. This
function satisfies $I_p(s, t) = 1-I_{1-p}(t, s)$.  Then 
\begin{align*}
R = \sum_{i=0}^{t-1} &{i+s-1 \choose i} p^s (1-p)^i +
\sum_{j=0}^{s-1}  {j+t-1 \choose j} p^j  (1-p)^t \\
   &= 1-I_p(s, t) + 1 - I_{1-p}(t, s) \\
   &= 1. 
\end{align*}
\end{proof}
Proving that \label{eqn:pmf} is a valid probability mass function.

@@

\begin{figure}[p!]
\begin{center}
\includegraphics[scale=.9]{shapes.pdf}
\end{center}
\caption{Different shapes of the SNB distribution with parameters ($s$, $t$, $p$), as given. \label{shapes.fig}}
\end{figure}

@@

For the special case of $\,s=t,$ the distribution of $\,Y\,$ is the riff-shuffle, or minimum negative binomial distribution~[\ref{Uppuluri 1970}].
Similar derivations of the closely-related maximum negative binomial discrete distributions also appear in~[\ref{Zhang 2000}, \ref{Zelterman 2004}].

@@

\section{The Compound Distribution}

@@

If we assume that $p$ has a Beta distribution, with constant prior parameters $\alpha$ and $\beta$ then we can calculate the closed form compound distribution of the SNB as
\begin{align}
\mathbb{P} \{Y = k | \mathbf{B}, \alpha, \beta \} &= {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} \{s \leq k \leq s+k-1\} + \nonumber \\
& {k-1 \choose k-t} \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} \{t \leq k \leq s+k-1\}
\end{align}
where $B$ is the Beta function and $\mathbf{B}$ is the sequence of Bernoulli data.

When $p$ is known, the probability mass function (PMF) of the number of steps until the process is stopped is given by (\ref{eqn:pmf}). However, many times this is not the case. In Phase II clinical trials for example patients are often enrolled sequentially until there are enough patients to guarantee the desired level of power or until a specified number of adverse events are recorded. In cases such as these the endpoints $s$ and $t$ are known and we would like to be able to estimate $p$ either during enrollment to provide odds of successful enrollment or after an endpoint has been reached to determine the amount of uncertainty in the endpoint. To do this, we will assume that $p$'s probability in the absence of available data is $\text{Beta}(1/2, 1/2)$, the Jeffrey's prior. This distribution was chosen for two reasons. First, it is the conjugate prior of the (Negative) Binomial distribution and updating the distribution as data are received is straightforward. Second, it is objective and limits the number of ``prior coin-flips'' to one. This means that data are more heavily influence the distributional characteristics of $p$ with more data providing more proportional influence.

\begin{prop}
The compound PMF of the Stopped Negative Binomial distribution with a Beta($\alpha$, $\beta$) prior is:
\begin{align}
f(k | s, t, \alpha, \beta) &= {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} \{s \leq k \leq s+k-1\} + \nonumber \\
& {k-1 \choose k-t} \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} \{t \leq k \leq s+k-1\}
\end{align}
where $B$ is the Beta function.
\end{prop}
\begin{proof}
For notational simplicity, assume that $s,t \leq k \leq s+k-1$. When this is not the case appropriate terms should be removed as dictated by the indicator functions.
\begin{align*}
f(k | s, t, \alpha, \beta) = \frac{1}{B(\alpha, \beta)} & \int_0^1 {k-1 \choose s-1} p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
 & {k-1 \choose k-t} p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp \\
= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} dp + \\
 & \frac{1}{B(\alpha, \beta)} {k-1 \choose k-t} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp
\end{align*}
The result follows by applying the definition of the Beta function to the integral terms.
\end{proof}

@@

Returning to the breast cancer trial introduced in the first section, a 
response rate of 2/12 in the first stage corresponds to an individual 
probability of responding of $p=0.1667$. The standard deviation of this 
probability is $\sqrt{p(1-p)/12}=0.1076$. If we were to use the Central 
Limit Theorem then the 95\% confidence interval around the mean would 
be $[-0.0442, 0.3776]$. While this approach is problematic since the number 
of samples (12 in this case) is small it does indicate how much uncertainty 
is built into the first stage. If we consider samples from both stages then 
the standard deviation drops to 0.0630 and the 95\% confidence interval 
would be $[0.0432, 0.2901]$.

@@

Properly quantifying uncertainty in the process implies quantifying 
uncertainty in $p$, which stems from two sources. First is the number of 
samples we receive. An estimate of $p$ will be based on data and generally 
the more data we have, the better our estimate. If the data are small, as 
they often are in clinical trials, the less reliable our estimate is. The 
second source comes from heterogeneity in the data. While models may assume 
that the probability of success $p$ is a single value this is often not the 
case in real-world data. In clinical trials for example an individual's chance 
of having an adverse outcome may be influenced by any number of underlying 
factors, which may vary across individual. In these cases $p$ may not be a 
single numeric value and may be better conceptualized as a realization over 
an entire distribution. A frequentest approach will not capture this variation.

@@

To better understand how the distributional estimate of $p$ changes in the number of steps needed for two successes data were simulated for each potential stopping point from two to 11. This procedure essentially estimates the shape parameters of a beta distribution. Fig. \ref{fig:simon} shows the change in the mode of the distribution, which is also the Maximum Likelihood Estimator (MLE) as well as the five percent quantile of the distribution. The graph shows, for example, that if a second success is found with the third enrollment then the MLE estimate of $p$ is 0.5 and with 95\% confidence $p$ is at least 0.25. In this case investigators could have moved on to the next stage of the trial with only three enrollments as opposed to 12.

@@

Table \ref{tab:simon} shows the actual values used to generate Fig. \ref{fig:simon}. In the beginning of this section it was shown that, at the end of stage two, the breast cancer study actually tests whether the probability of success is greater than 0.0432 with 95\% confidence and requires 35 patients to do so. With the Stopped Negative Binomial model we only need six patients for the same level of confidence. This represents a tremendous amount of savings in time, effort and resources.

@@

Fig \ref{fig:hypo} shows that at the eight enrollment the estimate of $p$ is 
about 0.19 and, from the lower 5\% quantile, we can see that there is
a 95\% probability that $p$ is greater than 2.5\%. 
Fig. \ref{fig:conditional_snb} shows the compound SNB distribution for this 
hypothetical trial at the eighth step.  Successes may occur 
at steps nine through 11 with probability 26.7\%. The remaining probability
mass corresponds to reaching the $11^{th}$ step without a second success.

@@

If there are two successful enrollments with the second one occurring 
on the $10^{th}$ trial the success probability $p$ is distributed as 
$Beta(2.5, 8.8)$. 
Using the mode, a point estimate yields $\hat{p} = 0.1667$. The standard 
deviation of the estimate is 0.1210. The probability of two successes during 
the trial is 0.6843 and the probability of failure is 0.3157. The distribution 
of the stopping times, along with contributions from two positive events and 
less than two positive events is shown below.

@@

We have presented a new discrete distribution by curtailed sampling rules common in early-stage clinical trials, which we refer to as the Stopped Negative Binomial distribution. The distribution models the stopping time of a sequential trial where the trial is stopped when a number of events are accumulated. The compound distribution was derived for the case when the event probability $p$ has a Beta distribution. Using a trial description from \url{clinicaltrials.gov} we showed how the SNB is an integral part of trial monitoring, and post-hoc analysis and can be used in a framework alternative to the Simon two-stage optimal design while providing better estimates of the event probability along with quantifying the uncertainty associated with the estimates. As a result, we were able to show fewer patient enrollments and achieve the same estimates, with added  in a hypothetical, but representative, clinical trial.

Current work focuses on the generalization of the distribution and the application in other areas of clinical trials. Adverse outcomes in particular are another area that could greatly benefit both from the presented distribution and its framework. For these trials monitoring may need to be performed not only on the outcome of an intervention but also on the safety of the trial. This especially in areas like late-stage cancer treatments where the treatment is harsh and patients may be forced to drop out as a result of the side-effects of the treatment; a matter independent of the outcome. Other application areas include providing design that allows clinicians to balance uncertainty and success probability with a minimum number of patient enrollees.

@@

>>>>

%                 Doubly-Stopped Binomial Distribution
%                        Kane and Zelterman
\typeout{}\typeout{}\typeout{}\typeout{}
\typeout{ Two-Stage, Phase II Designs }
\typeout{}\typeout{}
\typeout{ Michelle DeVeaux \& Michael Kane \& Dan Zelterman }
\typeout{}\typeout{}\typeout{}\typeout{}

\documentclass[12pt]{article}         %LaTeX preamble
\parskip   .4em
\jot  2ex
\parindent .3in
\textwidth      6.5in
\oddsidemargin    0in
\textheight 8.75in
\topmargin      -.5in
\baselineskip    26pt
\renewcommand{\baselinestretch}{1.5}
\renewcommand{\arraystretch}{.75}     % vertical spacing in tables
\usepackage{graphicx}                 % Needed to import pdf files
\usepackage{adjustbox}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{url}
\DeclareMathOperator*{\argmin}{argmin}
\newcommand*{\argminl}{\argmin\limits}
\begin{document}

@@

\vspace*{1.25in}
\begin{tabular}{l} 
 {\bf Michelle DeVeaux} \\
 {\bf Michael Kane${}^*$} \\
 {\bf Daniel Zelterman}  \\[.1in]

@@

In the Simon 2-stage Phase II clinical trial [\ref{Simon 1989}], a small 
number of patients are enrolled onto a therapeutic 
protocol for an initial stage. For each patient, we assess either success 
or failure according to a 
clinically useful criteria. Success might include tumor response to treatment
or perhaps that the response is maintained for a minimum specified duration.
In addition to failing to achieve the primary endpoint, treatment failure 
might include adverse reaction to the drug or withdrawal due to toxicity. We 
use a large number of these early successes as evidence for whether to treat
additional patients in a second stage. A statistically small number of 
successes in Stage I would lead us to terminate the trial without entering 
Stage II.

@@

% ---------------------------  2. Notation ----------------------
\section   {Notation and Summary of Distributional Results}
\label{notation.section}
% --------------------------------------------------------------

Let $\,B_1, B_2, \ldots \,$ denote a sequence of independent, identically 
distributed, Bernoulli random variables with $\mathbb{P}\{B_i=1\}=p$, for 
probability parameter $0\leq p \leq 1$. In the clinical trial setting 
$\,B_i = 0$ corresponds to a successful enrollment. 
Let $s$ and $t$ be positive integers. 
Define the stopped negative binomial (SNB) random variable $Y$ as the smallest 
integer value such that $\,\{B_1, \ldots , B_Y\}\,$ contains {\em either} 
$\,s\,$ successes or else $\,t\,$ failures. The distribution of $\,Y\,$ has 
support on integer values in the range 
\begin{equation*}                                     %   (3)
     \min(s,t) \leq \; Y \;\leq s+t-1  \label{range.y.eq}
\end{equation*}
and it is distributed as
\begin{equation} \label{eqn:pmf}
\mathbb{P}\{Y=k\} = S(k, p, s)  \{s \leq k\} + T(k, p, t) \{t \leq k \} 
\end{equation}
where
\begin{equation} \label{eqn:N}
S(k, p, s) = {k-1 \choose s-1} p^s (1-p)^{k-s} 
\end{equation}
is the negative binomial probability of $s$ successes and
\begin{equation} \label{eqn:R}
T(k, p, t) = {k-1 \choose k-t} p^{k-t} (1-p)^t
\end{equation}
is the negative binomial probability of $t$ failures. It should be noted that
this result expressed using the de Finetti where indicator functions and
sets are equivalent. A more complete explanation of this equivalence can be 
found in [\ref{Pollard 2002}].

@@

To illustrate that (\ref{eqn:pmf}) is the distribution of $Y$, consider the 
process $\mathbf{X} = \{X(k) : k \in \mathbb{Z}_{\ge 0} \}$ with $X(0)=0$ and
\begin{equation*} \label{eqn:proc}
X_{k+1} = X_k + B_{k+1} \left[ k-t < X_k < s \right]
\end{equation*}
where $B_{k+1}$ is once again distributed as Bernoulli($p$) for all $k$ in the domain of $\mathbf{X}$. The process can be conceptualized as a series of coin flips that stops when either $s$ heads or $t$ tails are reached. At each step a coin is flipped. If it is heads, the process advances one diagonally in the positive horizontal and vertical direction. Otherwise, it advances in the positive horizontal direction only. The process stops, $X_k$'s value does not change, after either $X_k = s$ or $X_k = k-t$. One example is given in Fig. \ref{fig:kane_viz}.

\begin{prop}
The distribution of the stopping time $\argminl_k [X_k \geq s, X_k \leq k-t]$ 
is equal to the SNB and is given in (\ref{eqn:pmf}). 
\end{prop}
\begin{proof}
The proof will proceed in two parts. First, a combinatorial justification will be given for the probability mass value on each element of the support. Second, it will be shown that the sum of the masses over the support sums to one.

@@

Let $I_p(s, t)$ be the {\em regularized incomplete beta function}, which is 
also one minus the c.d.f. of the negative binomial distribution. This
function satisfies $I_p(s, t) = 1-I_{1-p}(t, s)$.  Then 
\begin{align*}
R = \sum_{i=0}^{t-1} &{i+s-1 \choose i} p^s (1-p)^i +
\sum_{j=0}^{s-1}  {j+t-1 \choose j} p^j  (1-p)^t \\
   &= 1-I_p(s, t) + 1 - I_{1-p}(t, s) \\
   &= 1. 
\end{align*}
\end{proof}
Proving that \ref{eqn:pmf} is a valid probability mass function.

@@

\begin{figure}[p!]
\begin{center}
\includegraphics[width=\textwidth]{shapes.pdf}
\end{center}
\caption{Different shapes of the SNB distribution with parameters ($s$, $t$, $p$), as given. Red indicates mass contributed by hitting $s$, teal indicates
mass contributed by hitting $t$. \label{shapes.fig}}
\end{figure}

@@

For the special case of $\,s=t,$ the distribution of $\,Y\,$ is the 
riff-shuffle, or minimum negative binomial distribution~[\ref{Uppuluri 1970}].
Similar derivations of the closely-related maximum negative binomial discrete 
distributions also appear in~[\ref{Zhang 2000}, \ref{Zelterman 2004}]. However,
where the maximum negative binomial is the smallest number of coin-flips 
to observe at least $c$ successes {\em and} $c$ failures, the snb give
the number of coin flips to observe {\em either} $s$ heads or $t$ tails.

@@

\section{The Posterior Distribution}

@@

If we assume that $p$ has a Beta distribution, with constant prior parameters $\alpha$ and $\beta$ then we can calculate the closed form posterior 
distribution of the SNB as
\begin{align}
\mathbb{P} \{Y = k | \mathbf{B}, \alpha, \beta \} &= {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} \{s \leq k \leq s+k-1\} + \nonumber \\
& {k-1 \choose k-t} \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} \{t \leq k \leq s+k-1\}
\end{align}
where $B$ is the Beta function and $\mathbf{B}$ is the sequence of Bernoulli data.

When $p$ is known, the probability mass function (PMF) of the number of steps 
until the process is stopped is given by (\ref{eqn:pmf}). However, many times 
this is not the case. In Phase II clinical trials for example patients are 
often enrolled sequentially until there are enough patients to guarantee the 
desired level of power or until a specified number of adverse events are 
recorded. In cases such as these the endpoints $s$ and $t$ are known and we 
would like to be able to estimate $p$ either during enrollment to provide odds 
of successful enrollment or after an endpoint has been reached to determine 
the amount of uncertainty in the endpoint. 
%To do this, we will assume that $p$'s probability in the absence of 
%available data is $\text{Beta}(1/2, 1/2)$, the Jeffrey's prior. This 
%distribution was chosen for two reasons. First, it is the conjugate prior 
%of the (Negative) Binomial distribution and updating the distribution as data 
%are received is straightforward. Second, it is objective and limits the number 
%of ``prior coin-flips'' to one. This means that data are more heavily 
%influence the distributional characteristics of $p$ with more data providing 
%more proportional influence.

\begin{prop}
The posterior PMF of the Stopped Negative Binomial distribution with a Beta($\alpha$, $\beta$) prior is:
\begin{align}
f(k | s, t, \alpha, \beta) &= {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} \{s \leq k \leq s+k-1\} + \nonumber \\
& {k-1 \choose k-t} \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} \{t \leq k \leq s+k-1\}
\end{align}
where $B$ is the Beta function.
\end{prop}
\begin{proof}
For notational simplicity, assume that $s,t \leq k \leq s+k-1$. When this is not the case appropriate terms should be removed as dictated by the indicator functions.
\begin{align*}
f(k | s, t, \alpha, \beta) = \frac{1}{B(\alpha, \beta)} & \int_0^1 {k-1 \choose s-1} p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
 & {k-1 \choose k-t} p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp \\
= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} dp + \\
 & \frac{1}{B(\alpha, \beta)} {k-1 \choose k-t} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp
\end{align*}
The result follows by applying the definition of the Beta function to the integral terms.
\end{proof}

@@

Returning to the breast cancer trial introduced in the first section, consider
the first stage, having a sampel size of 12. A
response rate of 2/12 stage corresponds to an individual 
probability of responding of $p=0.1667$. The standard deviation of this 
probability is $\sqrt{p(1-p)/12}=0.1076$. If we were to use the Central 
Limit Theorem then the 95\% confidence interval around the mean would 
be $[-0.0442, 0.3776]$. While this approach is problematic since the number 
of samples (12 in this case) is small it does indicate how much uncertainty 
is built into the first stage. If we consider samples from both stages then 
the standard deviation drops to 0.0630 and the 95\% confidence interval 
would be $[0.0432, 0.2901]$.

@@

Properly quantifying uncertainty in the process implies quantifying 
uncertainty in $p$, which stems from two sources. First is the number of 
samples we receive. An estimate of $p$ will be based on data and generally 
the more data we have, the better our estimate. If the data are small, as 
they often are in clinical trials, the less reliable our estimate is. The 
second source comes from heterogeneity in the data. While models may assume 
that the probability of success $p$ is a single value this is often not the 
case in real-world data. In clinical trials for example an individual's chance 
of having an adverse outcome may be influenced by any number of underlying 
factors, which may vary across individual. In these cases $p$ may not be a 
single numeric value and may be better conceptualized as a realization over 
an entire distribution. A frequentist approach will not capture this variation.

@@

To better understand how the distributional estimate of $p$ changes in the 
number of steps needed for two successes data were simulated for each 
potential stopping point from two to 11. That is, we will fit the
Beta distribution with the Jeffrey's prior to the sequence of coin flips 
$\{[1,1], [1, 0, 1], [1, 0, 0, 1], ... \}$ in the support of the SNB with
$s=2$ and $t=11$.  Fig. \ref{fig:simon} shows the 
change in the mode of the beta distribution, which is also the Maximum 
Likelihood Estimator (MLE) as well as the five percent quantile of the 
distribution.  The graph shows, for example, that if a second success is 
found with the third enrollment then the MLE estimate of $p$ is 0.5 and with 
95\% confidence $p$ is at least 0.25. In this case investigators could have 
moved on to the next stage of the trial with only three enrollments as 
opposed to 12.

@@

Table \ref{tab:simon} shows the actual values used to generate Fig. 
\ref{fig:simon}. 
%In the beginning of this section it was shown that, at the 
%end of stage two, the breast cancer study actually tests whether the 
%probability of success is greater than 0.0432 with 95\% confidence and 
%requires 35 patients to do so. With the Stopped Negative Binomial model we 
%only need six patients for the same level of confidence. This represents a 
%tremendous amount of savings in time, effort and resources.

@@

Fig \ref{fig:hypo} shows that at the eighth enrollment the estimate of $p$ is 
about 0.19 and, from the lower 5\% quantile, we can see that there is
a 95\% probability that $p$ is greater than 2.5\%. 
Fig. \ref{fig:conditional_snb} shows the posterior SNB distribution for this 
hypothetical trial at the eighth step.  Successes may occur 
at steps nine through 11 with probability 26.7\%. The remaining probability
mass corresponds to reaching the $11^{th}$ step without a second success.

@@

If there are two successful enrollments with the second one occurring 
on the $10^{th}$ trial the success probability $p$ is distributed as 
$Beta(2.5, 8.5)$ (assuming the Jeffrey's prior). 
Using the mode, a point estimate yields $\hat{p} = 0.1667$. The standard 
deviation of the estimate is 0.1210. The probability of two successes during 
the trial is 0.6843 and the probability of failure is 0.3157. The distribution 
of the stopping times, along with contributions from two positive events and 
less than two positive events is shown below.

@@

We have presented a new discrete distribution by curtailed sampling rules common in early-stage clinical trials, which we refer to as the Stopped Negative Binomial distribution. The distribution models the stopping time of a sequential trial where the trial is stopped when a number of events are accumulated. The 
posterior distribution was derived for the case when the event probability $p$ 
has a Beta distribution. Using a trial description from 
\url{clinicaltrials.gov} we showed how the SNB is an integral part of trial 
monitoring, and post-hoc analysis and can be used in a framework alternative 
to the Simon two-stage optimal design while providing better estimates of the 
event probability along with quantifying the uncertainty associated with the 
estimates. As a result, we were able to show fewer patient enrollments and 
achieve the same estimates, when added in a hypothetical, but representative, 
clinical trial.

Current work focuses on the generalization of the distribution and the application in other areas of clinical trials. Adverse outcomes in particular are another area that could greatly benefit both from the presented distribution and its framework. For these trials monitoring may need to be performed not only on the outcome of an intervention but also on the safety of the trial. This especially 
true in areas like late-stage cancer treatments where the treatment is harsh 
and patients may be forced to drop out as a result of the side-effects of the 
treatment; a matter independent of the outcome. Other application areas 
include providing designs that allows clinicians to balance uncertainty and 
success probability with a minimum number of patient enrollees.

@@

\item[\bibref{Pollard 2002}]
Pollard D. {\it A User's Guide to Measure Theoretic Probability}, 
Cambridge University Press, 2002.
